---
title: "Refresher/crash course in basic R and tidyverse"
author: "Chong Kwek Yan"
date: "2024-12-17"
output: html_document
---

# Really basic basics

R as a calculator:

```{r}
1 + 2
```

```{r}
(1 - 2 * 3) / 4
```

```{r}
5 ^ 2
```

You can press the up arrow to go back again and edit the previous commands.

However, a good habit is to start a .R script file every time especially when you know you are going to be writing many commands.

Use `<-` or `->` to assign the output to an 'object'.

```{r}
a <- 12345
b <- 54321
c <- a + b

my.calculation <- c / a + b
```

You can then recall the object `my.calculation`.

If you assign the output to an object, there will be no feedback output in the console.

Spaces between operators don't matter, but note that R is caps-sensitive.

```{r}
My.Calculation <- a/(c+b)

my.calculation; My.Calculation
```

A 'function' runs a scripted bunch of calculations given 'arguments' which are the instructions:

```{r}
sum(a, b, c)
```

```{r}
sqrt(25)
```

$e^x$ with Euler's number, $e$, as the base:

```{r}
exp(2)
```

Logarithm with base 10:

```{r}
log10(a)
```

Logarithm specifying base 2:

```{r}
log(a, base = 2)
```

Otherwise the default base is Euler's number, i.e., $ln$). To find out how to use a function, what are the arguments, what are the default values for the arguments, etc., call up the help page by typing `?log`.

To find out which function to use to do something, just [Google](www.google.com)!

# Types of objects

A vector is a series of elements.

```{r}
c(a, b, c, 11, 22, 33)
```

Some functions are helpful for generating vectors:
```{r}
50:70
```

```{r}
seq(from = 1, to = 10, by = 0.1)
```

Note that you can omit the names of arguments if you specify them according to their default sequence in the function. But if not, e.g., if you skip an argument, then you must specify the argument name.

```{r}
seq(1, 10, length.out = 10)
```

```{r}
rep(NA, 10)
```

You can string up vectors in sequence:

```{r}
c(rep(1, 3), rep(2,3))
```

You can store vectors as objects and then perform calculations:

```{r}
y <- c(a, b, c, 11, 22, 33)
x1 <- c(rep(1, 3), rep(2,3))

y+x1
```

But not all vectors must consist of numbers.

```{r}
class(y)
```

```{r}
z <- c("hello", "how", "are", "you", "thank", "you")

class(z)
```

Data usually looks more familiar in the form of a table:

```{r}
eg.matrix <- cbind(y, x1)

eg.matrix
```

Note that a "matrix" in R means a table where all the elements are of the same 'class', e.g., `numeric` or `character`. If you have a mix of columns of different classes, you will be making a `data.frame`.

```{r}
eg.data <- data.frame(y, z)

eg.data
```

```{r}
class(eg.matrix); class(eg.data)
```

You can use square brackets to select specific rows, columns, and cells in the matrix or data frame.

```{r}
eg.matrix[1,]
```

```{r}
eg.matrix[,2]
```

```{r}
eg.matrix[1,2]
```

You can also use the header name to select columns.

For matrices,
```{r}
eg.data[,"y"]
```

For data frames, there are more options

```{r}
eg.data$y
```

```{r}
eg.data["y"]
```

To bind tables by columns:

```{r}
cbind(eg.matrix, eg.data)
```

But this only works if there are equal numbers of rows in the tables.

Combining rows:

```{r}
rbind(eg.matrix, eg.matrix)
```

But this only works if there are equal number of columns *and* the column names are the same.

```{r eval=FALSE}
rbind(eg.matrix, eg.data)
```

You can also directly a new column by creating it using a name.
```{r}
eg.data$x1 <- eg.matrix[,"x1"]

eg.data
```

Matrices or data frames must be "rectangular", i.e., each column has the same number of elements.

If you have a bunch of very different things, e.g., vectors of different lengths, or several tables, you can still group them together within a single object, called a `list`.

```{r}
eg.list <- list(eg.data, eg.matrix)
```

Lists can be very complicated, e.g., lists nested within lists.

You can name each part of a list, e.g., at the point of making the list:

```{r}
eg.list <- list(eg.data = eg.data, eg.matrix = eg.matrix)
```

You can then extract the parts by calling the name with the `$` sign:

```{r}
eg.list$eg.data
```

Or you can assign the names later. Here I'm showing how you can use names with spaces in it. But to call such parts of objects with spaces, you have to enclose them within `.

```{r}
names(eg.list) <- c("example data", "example matrix")

eg.list$`example data`
```

Only parts of objects, e.g., columns of data frames or a list element, can have spaces in the names. Objects in the namespace of R cannot have spaces in the name.

# R projects and the .RProj file

Finding out or setting the working directory is a critical first step in using R. Most basic R courses will teach you to use `setwd()`. But more often than not, this sets what is called an "absolute path" which most likely only works on your computer while being at least slightly different on someone else's computer.

Alternatively, you may be taught to click Session > Set Working Directory... but if you share your stuff with someone else, you need to separately explain where you set this working directory.

There are better ways that facilitate collaborative work. One of these ways is to always create an R project instead, by File > New Project...

Download any one of [my public repositories](https://github.com/kwekiii) as an example, e.g., [RSharings](https://github.com/kwekiii/RSharings).

Double-click on the file with the .RProj extension to open up an R session where the working directory is automatically where the .RProj file is.

# The .csv file

One popular way to store data as tables is in the format of "comma separated variables", or a .csv file. Each cell in this table is separated by the cell on its right by a comma; empty cells are a consecutive series of commas. A new row is a new line. You can open such a file in any text editor (e.g., Notepad) to see this. The main advantage of a .csv file is its simplicity and so will likely be useable across different software and in the future.

```{r eval=FALSE}
palm <- read.csv("./data/palm.csv")
```

The "." means where the working directory is. Subfolders are then indicated by "/". Note that this is the forward slash, not the back slash. You can also use double back-slashes.

```{r echo=FALSE}
palm <- read.csv("../data/palm.csv")
```

One drawback of .csv files is that the contents of the cells cannot have any commas in them or they will be interpreted to be column breaks. That row may then have more columns that other rows and R will return an error trying to read the file.

See `?read.csv` for other arguments that can be used (some of which are listed under `read.table()`).

The contents of the .csv file is read as a `data.frame`, so you can call columns using the column names.

Aside from numeric or character, there are other classes of vectors.

```{r}
class(palm$seedling)
```

You can also create a logical vector using conditions:

```{r}
palm$position == "Edge"
```

```{r}
palm$seedling > 0
```

```{r}
palm$litter >= 1.0
```

```{r}
palm$forest != "Bedok_reservoir"
```

Logical vectors can then be used to filter for a subset rows (or columns), i.e., only those that are TRUE will be selected.

```{r}
palm[palm$forest != "Bedok_reservoir",]
```

# Functions and loops

(Skipped.)

# Libraries

Any R user can publish a 'package' of their functions. Such a package is called a 'library'. To install a package officially published on the R repository (called CRAN),

```{r eval=FALSE}
install.packages("novelforestSG")
install.packages("tidyverse")
```

To load the package:
```{r}
library(novelforestSG)
library(tidyverse)
```

The `novelforestSG` package loads an object, `novelforest_data`, that is actually a list of several tables.

```{r}
names(novelforest_data)
```

The `tidyverse` package actually contains many packages which have functions that allow a different way of doing things from 'base' R (i.e., the core R without any libraries). For example, if you had wanted to create a new column of the basal area of each stem from its DBH:

```{r eval=FALSE}
novelforest_data$trees$basalArea <- pi * (novelforest_data$trees$dbh_2011/2)^2
```

With `tidyverse`, you can use the mutate function:

```{r}
mutate(novelforest_data$trees, basalArea = pi * (dbh_2011/2)^2)
```

If you wanted to filter for certain rows,

```{r eval=FALSE}
filter(novelforest_data$trees, species == "Hevea brasiliensis")
```

If you want to select only certain columns,

```{r eval=FALSE}
select(novelforest_data$trees, -c("stem"))
# or just -stem if there's only one to exclude
```

You can also rename columns.

```{r}
rename(novelforest_data$trees, DBH = dbh_2011)
```

One of the features with `tidyverse` is the 'pipe', which allows you to chain several functions in series. The output of the earlier function is taken to be the first argument of the next function.

```{r}
novelforest_data$trees %>%
  mutate(basalArea = pi * (dbh_2011/2)^2) %>%
  filter(species == "Hevea brasiliensis") %>%
  group_by(plot) %>%
  summarise(totBasalArea_rubber = sum(basalArea))
```

The `group_by()` function is often followed by the `summarise()` that applies a particular set of calculations to each level in the group.

```{r}
novelforest_data$plots %>%
  group_by(plot) %>%
  summarise(canopy_mean = mean(canopy),
            litter_mean = mean(litter))
```

Another very useful function in `tidyverse` is `pivot_wider()` (and its reverse, `pivot_longer`).

```{r}
novelforest_data$pres %>%
  select(-patch) %>%
  distinct() %>%
  mutate(pres = 1) %>%
  pivot_wider(names_from = species,
              values_from = pres)
```

The last `tidyverse` functions to introduce are the 'joins' which merges tables together based on one or more shared columns.

```{r}
plot_sumRubber <- novelforest_data$trees %>%
  mutate(basalArea = pi * (dbh_2011/2)^2) %>%
  filter(species == "Hevea brasiliensis") %>%
  group_by(plot) %>%
  summarise(totBasalArea_rubber = sum(basalArea))

plot_meanEnv <- novelforest_data$plots %>%
  group_by(plot) %>%
  summarise(canopy_mean = mean(canopy),
            litter_mean = mean(litter))

full_join(plot_sumRubber, plot_meanEnv, by = "plot")
```

There are other joins, such as `left_join` or `inner_join`, that differ in terms of how rows that don't have a match in either of the tables are treated.

One of the best things about tidyverse is the use of pipes to chain up a series of functions so that only one object is produced.

```{r}
novelforest_data$trees %>%
  mutate(basalArea = pi * (dbh_2011/2)^2) %>%
  filter(species == "Hevea brasiliensis") %>%
  group_by(plot) %>%
  summarise(totBasalArea_rubber = sum(basalArea)) %>%
  full_join(novelforest_data$plots %>%
    group_by(plot) %>%
    summarise(canopy_mean = mean(canopy),
              litter_mean = mean(litter)),
    by = "plot")
```

# Save, load, and source

You can export tables using the `write.csv()` function. However, if you have several objects that you want to save together to be used together, or lists that cannot be exported as .csv tables, you can save an .RData file using save().

You can then load the .RData file using load(). When you do so, all the objects in the .RData file are loaded with the same names that they were saved with.

If you have a single object that you want to allow others to decide for themselves what names to use, you can save them with `saveRDS()` which saves an .RDS file.

# Last thing: R Markdown notebooks

This was written as an R Markdown notebook which allows text to be interspersed with 'chunks' of code.

One important thing about R notebooks is that the default working directory is **not** where the .RProj file is located; it is where the [.Rmd](https://github.com/kwekiii/RSharings/blob/master/overlookedBasics/overlookedBasics.Rmd) file is located. So you will have to specify the file paths from which to read or write data accordingly. Use `..` to go up one level in the file hierarchy.